Ten Simple Rules for Teaching Programming

### 2017-10-14 paper.tex
Changing affiliation
Addressing reviewers' comments

% To hide this, toggle the two macros below:
%\newcommand{\evidence}[1]{}
\newcommand{\evidence}[1]{\vspace{-0.3cm}\textbf{Evidence strength: #1}\vspace{0.5cm}}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten quick tips for teaching programming} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
{Neil C.C.\ Brown}\textsuperscript{1{\ddag}*},
{Greg~Wilson}\textsuperscript{2{\ddag}*}
\\
\bigskip
\textbf{1} King's College London / neil.c.c.brown@kcl.ac.uk \\
\textbf{2} Rangle.io / gvwilson@third-bit.com \\

@@

\rulemajor{There is no geek gene}{gene}
\evidence{Medium\footnote{Our evidence strength is an informal assessment, factoring in whether the effect has been replicated, and whether it has been shown specifically in computing, or only in other disciplines from which we believe it will transfer.}}

@@

\rulemajor{Use peer instruction}{peer-instruction}
\evidence{Very high}

@@

@@

A classic question in computing (and mathematics) education
is whether problems are better with context (e.g., find the highest student grade)
or without (e.g. find the maximum of the list of numbers).
Bouvier et al.~\cite{bouvier-context} examined this with a multi-university study
and found no difference between the two.
They suggest that since it makes no difference,
other considerations (such as motivation) should be primary.

@@

We offer one final observation: do not forget the human element.
Programmers have a reputation for pouring scorn on certain programming languages (e.g., PHP),
or for gatekeeping (e.g., stating that you cannot learn programming if you did not start young).
If you are teaching someone to program,
the last thing you want to do is make them feel like they can't succeed
or that any existing skill they have (no matter when or how acquired) is worthless.
Make your learners feel that they can be a programmer,
and they just might become one.

@@

@@

\begin{figure}[h]
\begin{mdframed}
\noindent
\textbf{Conventional Materials}
\begin{verbatim}
1. Click on "My Blocks" to see the blocks for components you created.
2. Click on "clap"
3. Drag out a when clap.Touched block
4. Click on “clapSound”
5. Drag out call clapSound.Play
6. Connect it after when clap.Touched
\end{verbatim}

>>>>

% To hide this, toggle the two macros below:
\newcommand{\evidence}[1]{}
%\newcommand{\evidence}[1]{\vspace{-0.3cm}\textbf{Evidence strength: #1}\vspace{0.5cm}}

@@

% Title must be 250 characters or less.
\begin{flushleft}
{\Large
\textbf\newline{Ten quick tips for teaching programming} % Please use "sentence case" for title and headings (capitalize only the first word in a title (or heading), the first word in a subtitle (or subheading), and any proper nouns).
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
{Neil C.C.\ Brown}\textsuperscript{1{\ddag}*},
{Greg~Wilson}\textsuperscript{2{\ddag}*}
\\
\bigskip
\textbf{1} King's College London / neil.c.c.brown@kcl.ac.uk \\
\textbf{2} DataCamp / gvwilson@third-bit.com \\

@@

These tips will be useful to anyone teaching programming at any level
and to any audience. A larger list aimed primarily at K-12 audiences
can be found at \cite{cs-teaching-tips}.

\rulemajor{Remember that there is no geek gene}{gene}
\evidence{Medium\footnote{Our evidence strength is an informal assessment, factoring in whether the effect has been replicated, and whether it has been shown specifically in computing, or only in other disciplines from which we believe it will transfer.}}

@@

\rulemajor{Use peer instruction}{peer-instruction}
\evidence{High}

@@

Note that live coding does not always have to start with a blank screen:
instructors may give students some starter code
that relies solely on concepts they have already mastered,
and then extend it or modify it with live coding.

@@

A classic question in computing (and mathematics) education
is whether problems are better with context (e.g., find the highest student grade)
or without (e.g. find the maximum of the list of numbers).
Bouvier et al.~\cite{bouvier-context} examined this with a multi-university study
and found no difference between the two.
They suggest that since it makes no difference,
other considerations (such as motivation) should be given priority.

@@

We offer one final observation: do not forget the human element.
Programmers have a reputation for pouring scorn on certain programming tools (e.g., pouring scorn on spreadsheets),
or for gatekeeping (e.g., stating that you cannot learn programming if you did not start young).
If you are teaching someone to program,
the last thing you want to do is make them feel like they can't succeed
or that any existing skill they have (no matter when or how acquired) is worthless.
Make your learners feel that they can be a programmer,
and they just might become one.

@@

\section*{Ten Quick Tips for Teaching Programming}

\begin{enumerate}
\item Remember that there is no geek gene.
\item Use peer instruction.
\item Use live coding.
\item Have students make predictions.
\item Use pair programming.
\item Use worked examples with labelled subgoals.
\item Stick to one language.
\item Use authentic tasks.
\item Remember that novices are not experts.
\item Don't just code.
\end{enumerate}

@@

\begin{figure}[h]
\begin{mdframed}
\noindent
\textbf{Conventional Materials}
\begin{verbatim}
1. Click on "My Blocks" to see the blocks for components you created.
2. Click on "clap"
3. Drag out a when clap.Touched block
4. Click on "clapSound"
5. Drag out call clapSound.Play
6. Connect it after when clap.Touched
\end{verbatim}
